The Sub-Millimetre User Reduction Facility: an iterative map-maker for SCUBA-2

### 2011-03-10 smurf_paper.tex
add description of NOI, and re-organize model component section

\subsubsection{Power spectral densities}

Next, we produce power spectral density (PSD) plots for several
representative bolometers in Fig.~\ref{fig:pspec}. To produce this
figure, we follow the convention that the PSD as a function of
frequency, $\mathbf{P}(f)$, is normalized such that the integral over
frequency gives the same variance as the time-series variance across
the full time series. In other words, given a bolometer signal
$\mathbf{b}(t)$,
%
\begin{equation}
\int \mathbf{P}(f) df = \langle\mathbf{b}^2(t)\rangle.
\end{equation}
%
However, we only show the PSD up to the Nyquist frequency, so the
conversion from variance measured in the frequency domain to the time
domain requires and additional multiplcation by 2; hence the units are
pW$^2$\,Hz$^{-1}$ rather than pW$^2$\,s.

@@

The subsequent components identified by the PCA analysis in
Fig.~\ref{fig:pca} vary wildly. Component 3, though not obvious from
its eigenvector plot, is almost a pure ramping waveform that only
appears in the s4a data, and produces the broad line featuress
centered over 45\,Hz and 70\,Hz in the 450\,\micron\ PSDs in
Fig.~\ref{fig:psd}. In other data sets, this third component has also
often instead exhibited a regularly-spaced series of spikes with
heights quantized into several amplitude families (again, only in the
s4a array). In both cases, the shape and phase of this signal is the
same in each bolometer, although the amplitude (and sign) varies, as
evidenced by the large random scatter in the s4a eigenvalue maps. It
is thought that these features are some sort of data acquisition
glitch as the spikes (when they appear) exhibit the expected point
response function of the anti-aliasing filter. Similarly, component 4
also appears to be a data acquisition glitch in the s4a array,
although along a single column, and with a peak at a lower frequency
of 40\,Hz, and a second feature right at the Nyquist frequency.

@@

In practice, the map-maker is highly configurable, both in terms of
the order in which model components are fit (and whether they are fit
at all), and the details of \emph{how} they are fit. In the following
sections we describe these details.

%-------------------------------------------------
\subsection{Model components}
%-------------------------------------------------

In addition to specifying when each model component is calculated
(normally an attempt is made to order them by decreasing amplitude in
the time-series), there are a number of parameters that control their
operation.

\subsection{\texttt{COM,GAI}: common-mode estimation}

@@

\subsection{\texttt{FLT}: fourier transform filter}

@@

\subsection{\texttt{SMO}: time-domain smoothing filter}

\subsection{\texttt{AST}: map estimation}

@@

\subsubsection{Map-based despiking}

In addition, the \texttt{AST} model can use the map and its variance
to provide additional despiking of the time-series. Unlike the
time-domain despiker described ???, this calculation uses the scatter
in the population of samples used to estimate the map pixel value
(from different times and bolometers) to reject outliers. This method
is more robust against false-positive detections of bright/compact
astronomical sources since transient features in the time-series are
unlikely to occur by chance whenever a bolometer crosses a specific
location on the sky, whereas real astronomical sources are at fixed
spatial locations.

\subsection{\texttt{NOI}: noise estimation}
\label{sec:noi}

@@

@@

\subsubsection{Time-domain de-spiking}

@@

\subsubsection{Step correction}

@@

\subsubsection{Gap filling / apodization}

@@

\subsubsection{Map-based despiking}

>>>>

\subsubsection{Power spectral densities}
\label{sec:psd}

Next, we produce power spectral density (PSD) plots for several
representative bolometers in Fig.~\ref{fig:pspec}. To produce this
figure, we follow the convention that the PSD as a function of
frequency, $\mathbf{P}(f)$, is normalized such that the integral over
frequency gives the same variance as the time-series variance across
the full time series. In other words, given a bolometer signal
$\mathbf{b}(t)$,
%
\begin{equation}
\label{eq:psd}
\langle\mathbf{b}^2(t)\rangle = \int \mathbf{P}(f) df .
\end{equation}
%
However, we only show the PSD up to the Nyquist frequency, so the
conversion from variance measured in the frequency domain to the time
domain requires and additional multiplcation by 2; hence the units are
pW$^2$\,Hz$^{-1}$ rather than pW$^2$\,s.

@@

The subsequent components identified by the PCA analysis in
Fig.~\ref{fig:pca} vary wildly. Component 3, though not obvious from
its eigenvector plot, is almost a pure ramping waveform that only
appears in the s4a data, and produces the broad line featuress
centered over 45\,Hz and 70\,Hz in the 450\,\micron\ PSDs in
Fig.~\ref{fig:pspec}. In other data sets, this third component has
also often instead exhibited a regularly-spaced series of spikes with
heights quantized into several amplitude families (again, only in the
s4a array). In both cases, the shape and phase of this signal is the
same in each bolometer, although the amplitude (and sign) varies, as
evidenced by the large random scatter in the s4a eigenvalue maps. It
is thought that these features are some sort of data acquisition
glitch as the spikes (when they appear) exhibit the expected point
response function of the anti-aliasing filter. Similarly, component 4
also appears to be a data acquisition glitch in the s4a array,
although along a single column, and with a peak at a lower frequency
of 40\,Hz, and a second feature right at the Nyquist frequency.

@@

In practice, the map-maker is highly configurable, both in terms of
the order in which model components are fit (and whether they are fit
at all), and the details of \emph{how} they are fit. A summary of the
model components is given in the next section.

%-------------------------------------------------
\subsection{Model components}
\label{sec:components}
%-------------------------------------------------

In addition to specifying when each model component is calculated
(normally an attempt is made to order them by decreasing amplitude in
the time-series), there are a number of parameters that control their
operation. Table~\ref{tab:components} list the model components and
the typical order in which they are calculated. The following sections
describe their implementation in detail.

\begin{table}
  \caption{Summary of the model components that can be fit to
    \scuba\ time-series data with SMURF. Only the first group of
    models are typically fit to the data (\texttt{COM}--\texttt{NOI})
    in the indicated order. The remaining models
    (\texttt{DKS}--\texttt{PLN}) have only been included for
    completeness}
  \vspace{0.2cm}
  \centering
  \begin{tabular}{c|l}
    \hline
    Model & Description \\
    \hline
    \texttt{COM} & remove common-mode signal \\
    \texttt{GAI} & common-mode scaled to each bolometer \\
    \texttt{EXT} & extinction correction \\
    \texttt{FLT} & Fourier transform filter \\
    \texttt{AST} & map estimate of astronomical signal \\
    \texttt{NOI} & noise estimation \\
    \hline
    \texttt{DKS} & dark squid cleaning along columns \\
    \texttt{SMO} & time-domain smoothing filter \\
    \texttt{PLN} & 2-dimensional time-varying plane removal \\
    \hline
    \end{tabular}
  \label{tab:components}
\end{table}

\subsubsection{\texttt{COM,GAI}: common-mode estimation}
\label{sec:comgai}

@@

\subsubsection{\texttt{EXT}: extinction correction}
\label{sec:ext}

AGG

This is actually only calculated once before the map solution begins
using external opacity monitors (WVM, $\tau_\mathrm{CSO}$,
skydip). Summarize the measurements, order of preference and something
about the calibration for the SCUBA-2 filters.

\subsubsection{\texttt{FLT}: fourier transform filter}
\label{sec:flt}

@@

\subsubsection{\texttt{AST}: map estimation}
\label{sec:ast}

@@

In addition to map estimation, the \texttt{AST} model can also be used
to perform map-based despiking of the time-series
(Section~\ref{sec:mapdespike}), and to apply constraints to the map to
improve convergences (Section~???).

\subsubsection{\texttt{NOI}: noise estimation}
\label{sec:noi}

@@

As in Section~\ref{sec:psd} the FFT of each bolometer time series is
taken, and the PSD calculated. An average white noise level is then
measured from 2 to 10\,Hz, a relatively clean region that lies above
the $1/f$ knee, but below the high-frequency line features for most of
the bolometer data taken during S2SRO (Fig.~\ref{fig:pspec}). Taking
this constant level for $\mathbf{P}(f)$ we then calculate the expected
variance of the time-series (in approximately 200\,Hz samples) using
Eq.~\ref{eq:psd}. If the bolometer noise were produced purely by
uncorrelated thermal sources (i.e. no other long-scale drifts), with
no high-frequency line features, and without the attenuation at even
higher frequencies by the anti-aliasing filter, this is the
theoretical noise limit of the detectors. These measured variances are
stored, and then used in subsequent iterations to weight the data
points when calculating the map estimate (Section~\ref{sec:ast}).

Since noise estimation is usually calculated as the final step, it is
the cleanest data for which some cleaning operations may be
performed. For this reason \texttt{NOI} can optionally use the
time-domain DC step fixer (Section~\ref{sec:steps}) and spike
detection (Section~\ref{sec:timedespike}).

\subsubsection{\texttt{DKS}: dark squid cleaning along columns}
\label{sec:dks}

Columns share significant magnetic field pickup which is measured by
the dark squids. This model has been included although it makes little
improvement, or injects noise since the magnetic field pickup noise is
generally sub-dominant.

\subsubsection{\texttt{SMO}: time-domain smoothing filter}
\label{sec:smo}

\subsubsection{\texttt{PLN}: 2d time-varying sky noise plane removal}
\label{sec:pln}

We can remove a plane at each timeslice but it doesn't seem to improve
things. Perhaps refer to other papers that claim to have seen coherent
2d structures in the sky noise?

@@

\subsubsection{Time-domain de-spiking}
\label{sec:timedespike}

@@

\subsubsection{Step correction}
\label{sec:steps}

@@

\subsubsection{Gap filling / apodization}
\label{sec:gaps}

@@

\subsubsection{Map-based despiking}
\label{sec:mapdespike}

It is possible to use the map estimate (Section~\ref{sec:ast}) to
provide additional despiking of the time-series. Unlike the
time-domain despiker (Section~\ref{sec:timedespike}), this calculation
uses the scatter in the population of samples used to estimate the map
pixel values (from different times and bolometers) to reject
outliers. This method is more robust against false-positive detections
of bright/compact astronomical sources since transient features in the
time-series are unlikely to occur by chance whenever a bolometer
crosses a specific location on the sky, whereas real astronomical
sources are at fixed spatial locations.
